IFC Inside: A General Approach to Retrofitting Languages with Dynamic Information Flow Control

### 2014-07-08 paper/real.tex
1st paragraph lio

\subsection{Haskell}
\label{sec:real:hs}
In contrast with the C embedding which relies on hardware protection
mechanisms, a Haskell implementation can leverage Haskell's strong data
abstraction and static type system, monadic approach to effects, and
lightweight concurrency to implement the embedding in a more lightweight
manner.  We briefly describe one such system, LIO~\cite{lio}, which
implements IFC as a library.

LIO is implemented by defining a new monad, \verb|LIO|, which wraps Haskell's \verb|IO|
monad.
%
The purpose of this monad is twofold: it restricts the use of
arbitrary effects that would ordinarily be allowed by the \verb|IO| monad,
and it associates labels with tasks.
%
Computations in the \verb|LIO|
monad can be thought to be operating within the IFC system.
%
One important aspect of using \verb|IO| as the base for this
implementation is that it allows use of Haskell's efficient
implementation of threads, channels, etc. (e.g., in the concurrent
version of LIO, we use Haskell's \texttt{forkIO} to fork a lightweight
thread in the case of |fork|~\cite{stefan:addressing-covert}), in
contrast to defining them in a completely pure fashion (as suggested
in Section~\ref{sec:monad}).

What is the interpretation of this system as per Section~\ref{sec:retrofit}?
%
Here, the \emph{pure subset} of Haskell is the target language, while
the monadic subset of Haskell in the \verb|LIO| monad is the IFC
language.
%
Crucially, the lack of unrestricted mutation in the pure fragment of
Haskell prevents direct communication between tasks, even when memory is
shared between them.\footnote{However, lazy evaluation can still produce
a covert channel.}

Since this IFC system is implemented as a library in Haskell,
implementation-wise, we must ensure that these languages are indeed the subsets of Haskell
we claim, i.e., while the concrete language is all of Haskell, we must
ensure that we can restrict programs to our subset of Haskell that
encodes the combined language.
%
To this end, \verb|LIO| relies on Haskell's strong data abstraction and type system
(as enforced by Safe Haskell~\cite{Terei:2012:SH:2364506.2364524}) to
ensure that arbitrary \verb|IO| actions cannot be lifted into
\verb|LIO|.
%
In other words, assuming \verb|LIO| is implemented correctly, programs
written in the \verb|LIO| monad cannot perform arbitrary \verb|IO| actions
without breaking abstraction.

We refer the interested reader to~\cite{lio,stefan:addressing-covert} for
additional details on the various implementations of this system.\footnote{It's worth noting that the proofs of non-interference we have given for asynchronous communication primitives are new and not in the original presentation of LIO.}

\Red{TODO: note that if the target language allows inspecting if
  result is exception we can implement Breeze's NAV (vs. just
propagating)}

>>>>

\subsection{Haskell}
\label{sec:real:hs}
Our work borrows many ideas from the Haskell coarse-grained IFC
system LIO~\cite{lio, stefan:addressing-covert}.
%
LIO relies on Haskell's strong type system and monadic encoding of
side-effecting computations to achieve isolation and define the IFC
sub-language.
%
Specifically LIO provides the \verb|LIO| monad as a way of restricting
(almost all) side-effects; this monad also keeps track of a current
label which is, in turn, used by functions such as |getLabel| to
define the IFC core.
%
In the context of our framework, LIO can be understood as follows: the
\emph{pure subset} of Haskell is the target language, while the
monadic subset of Haskell in the \verb|LIO| monad is the IFC language.

Indeed, by rephrasing LIO in terms of our framework, we simplified
LIO's treatment of exceptions, which have since been integrated in the
code-base.
%
we addressed a limitation in LIO.
